Implementing XBF: An Efficient Self-Describing Binary Format

### 2023-08-18 docs/presentation/presentation.tex
Add visual examples for inefficiencies
Update presentation

\title{Implementing XBF: An Efficient Self-Describing Binary Format}
% \subtitle{Using \LaTeX\ to prepare slides}
% \course{Master's Degree in Computer Science}
\author{\href{mailto:dkrautha@pm.me}{David Krauthamer}}
% \IDnumber{1234567}

@@

% possibly add three slides explaining each of these
% 1: XML being huge
% 2: plain text example from paper
% 3: 1027 as u32 0 4 3 0 -> 0 3 4 0

@@

\begin{frame}{Primitive Metadata}
	\center
	\begin{tabular}{|c|c|}
		\hline
		\textbf{Type} & \textbf{Discriminant} \\
		\hline
		Boolean       & 0                     \\
		U8            & 1                     \\
		U16           & 2                     \\
		U32           & 3                     \\
		...           & ...                   \\
		I256          & 12                    \\
		F32           & 13                    \\
		F64           & 14                    \\
		Bytes         & 15                    \\
		String        & 16                    \\
		\hline
	\end{tabular}
	% two columsn for each half of the list
	\note[item]{Metadata is sent as a single byte discriminant value}
	\note[item]{The discriminant begin at 0 with Boolean, and increase by 1 for each type}
	\note[item]{Progress in the same order as the previous slide, with String being the final value}
\end{frame}

@@

\begin{frame}[fragile]{Struct Metadata Example}
	\begin{columns}
		\begin{column}{0.65\textwidth}
			\note[item]{ DS = Discriminant of a Struct

				DX = Discriminant of field "first"

				DY = Discriminant of field "last"}
			\center
			\begin{bytefield}{32}
				\bitbox{2}{\rot{DS}} & \bitbox{16}{struct name length} & \bitboxes{2}{Point} & \bitbox{4}{fields} \\
				\bitboxes{2}{{18} {5} {0} {0} {0} {0} {0} {0} {0} {80} {\rot{111}} {\rot{105}} {\rot{110}} {\rot{116}} {2} {0}} \\
			\end{bytefield}
			\begin{bytefield}{32}
				\bitbox{16}{first name length} & \bitboxes{2}{first} & \bitbox{2}{\rot{DX}} \\
				\bitboxes{2}{{5} {0} {0} {0} {0} {0} {0} {0} {\rot{102}} {\rot{105}} {\rot{114}} {\rot{115}} {\rot{116}} {7}} \\
			\end{bytefield}
			\begin{bytefield}{32}
				\bitbox{16}{last name length} & \bitboxes{2}{last} & \bitbox{2}{\rot{DY}} \\
				\bitboxes{2}{{4} {0} {0} {0} {0} {0} {0} {0} {\rot{108}} {\rot{97}} {\rot{115}} {\rot{116}} {7}}
			\end{bytefield}
		\end{column}
		\begin{column}{0.35\textwidth}
			\begin{lstlisting}
struct Point {
	first: i8,
	last: i8
}
\end{lstlisting}
		\end{column}
	\end{columns}
	% add explanation for DS, DX, DY
\end{frame}

>>>>

\title{Implementing XBF: An Efficient Self-Describing Binary Format}
\author{\href{mailto:dkrautha@pm.me}{David Krauthamer}}

@@

\begin{frame}[fragile]{Repeated Metadata}
	\begin{columns}
		\begin{column}{0.5\textwidth}
			\note[item]{XML example where a huge percent of the actual data sent is metadata}
			\begin{lstlisting}[language=XML]
<person>
	<name>John Jackson</name>
	<age>25</age>
</person>
\end{lstlisting}
		\end{column}
		\begin{column}{0.5\textwidth}
			\begin{table}
				\caption{XML Space Utilization}
				\center
				\begin{tabular}{|c|c|c|}
					\hline
					\textbf{} & \textbf{Bytes Used} & \textbf{Percent} \\
					\hline
					Metadata  & 45                  & 76\%             \\
					Data      & 14                  & 24\%             \\
					\hline
					Total     & 59                  & 100              \\
					\hline
				\end{tabular}
			\end{table}
		\end{column}
	\end{columns}
\end{frame}

\begin{frame}[fragile]{Plain Text Example}
	\begin{table}
		\caption{Sending A 32-bit Unsigned Integer}
		\begin{center}
			\begin{tabular}{|c|c|c|}
				\hline
				\textbf{Number} & \textbf{ASCII Bytes} & \textbf{Binary Bytes} \\
				\hline
				1               & 1                    & 4                     \\
				9999            & 4                    & 4                     \\
				7654321         & 7                    & 4                     \\
				\hline
			\end{tabular}
		\end{center}
	\end{table}
\end{frame}

\begin{frame}[fragile]{Big Endian Example: 32-bit Unsigned Integer}
	\begin{center}
		\textbf{Number To Send: 1027} \\
	\end{center}
	\begin{columns}
		\begin{column}{0.3\textwidth}
			\begin{bytefield}{16}
				\wordbox{1}{Sender (Little Endian)} \\
				\bitboxes{4}{{3} {4} {0} {0}} \\
			\end{bytefield}
		\end{column}
		\begin{column}{0.3\textwidth}
			\begin{bytefield}{16}
				\wordbox{1}{On the Wire (Big Endian)} \\
				\bitboxes{4}{{0} {0} {4} {3}} \\
			\end{bytefield}
		\end{column}
		\begin{column}{0.3\textwidth}
			\begin{bytefield}{16}
				\wordbox{1}{Recipient (Little Endian)} \\
				\bitboxes{4}{{3} {4} {0} {0}} \\
			\end{bytefield}
		\end{column}
	\end{columns}
\end{frame}

@@

\begin{frame}{Primitive Metadata}
	\begin{columns}
		\begin{column}{0.5\textwidth}
			\center
			\begin{tabular}{|c|c|}
				\hline
				\textbf{Type} & \textbf{Discriminant} \\
				\hline
				Boolean       & 0                     \\
				U8            & 1                     \\
				U16           & 2                     \\
				U32           & 3                     \\
				U64           & 4                     \\
				U128          & 5                     \\
				U256          & 6                     \\
				\hline
			\end{tabular}
		\end{column}
		\begin{column}{0.5\textwidth}
			\center
			\begin{tabular}{|c|c|}
				\hline
				\textbf{Type} & \textbf{Discriminant} \\
				\hline
				I8            & 7                     \\
				I16           & 8                     \\
				I32           & 9                     \\
				I64           & 10                    \\
				I128          & 11                    \\
				I256          & 12                    \\
				F32           & 13                    \\
				F64           & 14                    \\
				Bytes         & 15                    \\
				String        & 16                    \\
				\hline
			\end{tabular}
		\end{column}
	\end{columns}
	\note[item]{Metadata is sent as a single byte discriminant value}
	\note[item]{The discriminant begin at 0 with Boolean, and increase by 1 for each type}
\end{frame}

@@

\begin{frame}[fragile]{Struct Metadata Example}
	\begin{columns}
		\begin{column}{0.65\textwidth}
			\center
			\begin{bytefield}{32}
				\bitbox{2}{\rot{DS}} & \bitbox{16}{struct name length} & \bitboxes{2}{Point} & \bitbox{4}{fields} \\
				\bitboxes{2}{{18} {5} {0} {0} {0} {0} {0} {0} {0} {80} {\rot{111}} {\rot{105}} {\rot{110}} {\rot{116}} {2} {0}} \\
			\end{bytefield}
			\begin{bytefield}{32}
				\bitbox{16}{first name length} & \bitboxes{2}{first} & \bitbox{2}{\rot{DF}} \\
				\bitboxes{2}{{5} {0} {0} {0} {0} {0} {0} {0} {\rot{102}} {\rot{105}} {\rot{114}} {\rot{115}} {\rot{116}} {7}} \\
			\end{bytefield}
			\begin{bytefield}{32}
				\bitbox{16}{last name length} & \bitboxes{2}{last} & \bitbox{2}{\rot{DL}} \\
				\bitboxes{2}{{4} {0} {0} {0} {0} {0} {0} {0} {\rot{108}} {\rot{97}} {\rot{115}} {\rot{116}} {7}}
			\end{bytefield}
		\end{column}
		\begin{column}{0.35\textwidth}
			\begin{lstlisting}
struct Point {
	first: i8,
	last: i8
}
\end{lstlisting}
			\hfill \break
			DS = Struct Discriminant

			DF = Field "first" discriminant

			DL = Field "last" discriminant
		\end{column}
	\end{columns}
\end{frame}
